---
title: "Assignment1 - 607"
author: Sangeetha Sasikumar
Date: September 3, 2022
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```
Introduction:
The article I chose to do my analysis on wasn't an actual pure "news article". "Club Soccer Predictions": https://projects.fivethirtyeight.com/soccer-predictions/ was more of a listing of predictions and current rankings. 
I chose this dataset because the World Cup is around the corner and I am a huge fan of soccer. 

```{r}
soccer_data<-read.csv('https://raw.githubusercontent.com/Sangeetha-007/R-Practice/master/607/Assignments/Assignment1/spi_global_rankings_intl.csv')
head(soccer_data)
```

Summary of the International Soccer Power Index data
```{r, echo=FALSE}
summary(soccer_data)

```

Created a subset with the 10 team international teams
```{r}
soccer_subset<-head(soccer_data, 10)
soccer_subset
```

Renamed the columns of the subset
```{r}
colnames(soccer_subset) <- c("Ranking", "Country", "Association", "Offense", "Defense", "Soccer Power Index")
head(soccer_subset)

```
Looked at the summary of the new subset
```{r}
summary(soccer_subset)
```

Created a new csv of the subset so it can be uploaded to github
```{r}
write.csv(soccer_subset, file="SoccerSubset.csv", row.names=FALSE)
getwd()

```
Bar graph on Defense per Country
```{r}
ggplot(soccer_subset, aes(x=Country, y=Defense)) + geom_bar(stat="identity") + 
  labs(x="Country", y="Defense") +ggtitle("International Soccer Team Defense Values")

```

Bar graph on Offense per Country
```{r}
ggplot(soccer_subset, aes(x=Country, y=Offense)) + geom_bar(stat="identity") + 
  labs(x="Country", y="Offense") +ggtitle("International Soccer Team Offense Values")
```


The country with the highest defense is Belgium, while the country with the highest offense is Germany. This is interesting to me
because I didn't think Belgium is a strong soccer team at all. I would have assumed Argentina or Brazil would have higher values. 

Conclusion/Findings and Recommendations:

From my analysis, it was interesting to see how highest defense is Belgium and highest offense is Germany. I can't make further
analysis based on the data presented in the zip file to take this analysis further, but if I find supporting data elsewhere, it may
be beneficial. Data based on the number of fans or population for each national team would be helpful and I can probably check if there
is a correlation to the offenses and defenses.
