---
title: 'Inference for numerical data'
author: "Sangeetha Sasikumar"
date: "10/20/2022"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```

## Getting Started

### Load packages

In this lab, we will explore and visualize the data using the **tidyverse** suite of packages, and perform statistical inference using **infer**. The data can be found in the companion package for OpenIntro resources, **openintro**.

Let's load the packages.

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(infer)
library(ggplot2)
library(psych)
```


### The data

Every two years, the Centers for Disease Control and Prevention conduct the Youth Risk Behavior Surveillance System (YRBSS) survey, where it takes data from high schoolers (9th through 12th grade), to analyze health patterns. You will work with a selected group of variables from a random sample of observations during one of the years the YRBSS was conducted.

Load the `yrbss` data set into your workspace.

```{r load-data}
data('yrbss', package='openintro')
```

There are observations on 13 different variables, some categorical and some numerical. The meaning of each variable can be found by bringing up the help file:

```{r help-nc, eval=FALSE}
?yrbss
```


1.  What are the cases in this data set? How many cases are there in our sample?

Remember that you can answer this question by viewing the data in the data viewer or by using the following command:

```{r str}
glimpse(yrbss)

nrow(yrbss)
```
I checked using glimpse and nrow to determine there are 13,583 rows or cases. 

## Exploratory data analysis

You will first start with analyzing the weight of the participants in kilograms: `weight`.

Using visualization and summary statistics, describe the distribution of weights. The `summary` function can be useful.

```{r summary}
summary(yrbss$weight)
```

2.  How many observations are we missing weights from?
We have NA values for 1,004 obeservations, that's how much we have missing weights from. 

Next, consider the possible relationship between a high schooler's weight and their physical activity. Plotting the data is a useful first step because it helps us quickly visualize trends, identify strong associations, and develop research questions.

First, let's create a new variable `physical_3plus`, which will be coded as either "yes" if they are physically active for at least 3 days a week, and "no" if not.

```{r create new var}
yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(yrbss$physically_active_7d > 2, "yes", "no"))

yrbss

```




3.  Make a side-by-side boxplot of `physical_3plus` and `weight`. Is there a relationship between these two variables? What did you expect and why?
```{r}
sum(is.na(yrbss$physical_3plus))

ggplot(yrbss, aes(x=weight, y=physical_3plus)) + geom_boxplot() + theme_dark()
```
The relationship between a student’s weight and whether they are physically active at least 3 times per week from the graph shows how those who are not physically active at least 3 times per week weigh less than those who are physically active at least 3 times per week. This is interesting as we would have expected those who were physically active to weigh less than those who were not physically active at least 3 times per week. This can be due to the fact since they are all teenagers and are still growing, they need a lot of food. Growing plus being physically active will require even more food. Also as someone who trains at the gym, I know muscle weighs more than fat!

The box plots show how the medians of the two distributions compare, but we can also compare the means of the distributions using the following to first group the data by the `physical_3plus` variable, and then calculate the mean `weight` in these groups using the `mean` function while ignoring missing values by setting the `na.rm` argument to `TRUE`.

```{r by-means}
yrbss %>%
  group_by(physical_3plus) %>%
  summarise(mean_weight = mean(weight, na.rm = TRUE))
```

There is an observed difference, but is this difference statistically significant? In order to answer this question we will conduct a hypothesis test.

## Inference

4.  Are all conditions necessary for inference satisfied? Comment on each. You can compute the group sizes with the `summarize` command above by defining a new variable with the definition `n()`.
According to Khan Academy, the conditions we need for inference on a mean are:
1. Random: A random sample or randomized experiment should be used to obtain the data.
2. Normal: The sampling distribution of \bar x,(the sample mean) needs to be approximately normal. This is true if our parent population is normal or if our sample is reasonably large (n≥30). 
3. Independent: Individual observations need to be independent. If sampling without replacement, our sample size shouldn't be more than 10\%10%10, percent of the population.
According to the data, we can see that it is a representative sample of many students across national, state,and tribal territories. All students are independent of each other. According to the box plots above, we can also see that the data appears to be normally distributed. However, we did not look at a random sample. 
```{r}

yrbss %>% 
  group_by(physical_3plus) %>% 
  summarise(freq = table(weight)) %>%
  summarise(n = sum(freq))
```

5.  Write the hypotheses for testing if the average weights are different for those who exercise at least times a week and those who don't.

Next, we will introduce a new function, `hypothesize`, that falls into the `infer` workflow. You will use this method for conducting hypothesis tests. 

But first, we need to initialize the test, which we will save as `obs_diff`.

```{r, warning = FALSE}
#obs_diff <- yrbss %>%
#  specify(weight ~ physical_3plus) %>%
#  calculate(stat = "diff in means", order = c("yes", "no"))
```

Notice how you can use the functions `specify` and `calculate` again like you did for calculating confidence intervals. Here, though, the statistic you are searching for is the difference in means, with the order being `yes - no != 0`.

After you have initialized the test, you need to simulate the test on the null distribution, which we will save as `null`.

```{r inf-weight-habit-ht-null, tidy=FALSE, warning = FALSE}
null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```

Here, `hypothesize` is used to set the null hypothesis as a test for independence. In one sample cases, the `null` argument can be set to "point" to test a hypothesis relative to a point estimate.

Also, note that the `type` argument within `generate` is set to `permute`, whichis the argument when generating a null distribution for a hypothesis test.

We can visualize this null distribution with the following code:

```{r}
ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()
```


6. How many of these `null` permutations have a difference of at least `obs_stat`?

Now that the test is initialized and the null distribution formed, you can calculate the p-value for your hypothesis test using the function `get_p_value`.
```{r}
yrbss <- yrbss %>% 
  mutate(physical_3plus = if_else(physically_active_7d > 2, "yes", "no"))
yrbss %>%
  group_by(physical_3plus) %>%
  summarise(mean_weight = mean(weight, na.rm = TRUE))   
```

```{r inf-weight-habit-ht-pvalue}
null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")
```

This the standard workflow for performing hypothesis tests.

7.  Construct and record a confidence interval for the difference between the weights of those who exercise at least three times a week and those who don't, and interpret this interval in context of the data.
```{r}
yrbss %>% 
  group_by(physical_3plus) %>% 
  summarise(sd_weight = sd(weight, na.rm = TRUE))


```
The standard deviation is 17.6 for those who do are not physically active at least 3 days per week and 16.5 for those who are.

```{r}
#means of the weights
yrbss %>% 
  group_by(physical_3plus) %>% 
  summarise(mean_weight = mean(weight, na.rm = TRUE))

#sample size for each category
yrbss %>% 
  group_by(physical_3plus) %>% 
  summarise(freq = table(weight)) %>%
  summarise(n = sum(freq))

xnot3 <- 66.67389
nnot3 <- 4022
snot3 <- 17.63805
x3 <- 68.44847
n3 <- 8342
s3 <- 16.47832

z = 1.96

uci_not <- xnot3 + z*(snot3/sqrt(nnot3))
lci_not <- xnot3 - z*(snot3/sqrt(nnot3))
uci_not

lci_not

uci <- x3 + z*(s3/sqrt(n3))
lci <- x3 - z*(s3/sqrt(n3))

uci
lci
```
There's 95% confidence that students who exercise at least three times a week have an average weight between 68.09 kg and 68.8 kg. Also those students who do not exercise at least three times a week have an average weight between 66.13 kg and 67.22 kg with 95% confidence. 

## More Practice

8.  Calculate a 95% confidence interval for the average height in meters (`height`) and interpret it in context.
```{r}
tb <- as.data.frame(table(yrbss$height))
freq <- sum(tb$Freq)

mean_height <- mean(yrbss$height, na.rm = TRUE)
sd_height <- sd(yrbss$height, na.rm = TRUE)
sample_height <- yrbss %>% 
  summarise(freq = table(height)) %>%
  summarise(n = sum(freq, na.rm = TRUE))

height_upper <- mean_height + z*(sd_height/sqrt(sample_height))

height_lower <- mean_height - z*(sd_height/sqrt(sample_height))

c(height_lower,height_upper)

```
From a 95% CI, we can say the average height of the students of the population is between 1.68 m and 1.69 m.

9.  Calculate a new confidence interval for the same parameter at the 90% confidence level. Comment on the width of this interval versus the one obtained in the previous exercise.

Greater lower-bound value yet smaller upper bound value for CI at 90% confidence level. Taking the difference of the two confidence intervals from 90% and 95%, we see they have the same difference from each other. Therefore, the width of these two intervals is the same.
```{r}
z2 <- 1.645

height_upper_2 <- mean_height + z2*(sd_height/sqrt(sample_height))

height_lower_2 <- mean_height - z2*(sd_height/sqrt(sample_height))

c(height_lower_2 ,height_upper_2)

x <- abs(height_lower_2 - height_lower)
y <- abs(height_upper_2 - height_upper)

c(x,y)

```

10.  Conduct a hypothesis test evaluating whether the average height is different for those who exercise at least three times a week and those who don't.
```{r}
obs_diff_hgt <- yrbss %>%
  specify(height ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
set.seed(45698)
null_dist_hgt <- yrbss %>%
  specify(height ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
visualize(null_dist_hgt) + 
  shade_p_value(obs_stat = obs_diff_hgt, direction = "two_sided")
null_dist_hgt %>%
  get_p_value(obs_stat = obs_diff_hgt, direction = "two_sided")

# Non active
mean_height_2 <- 1.6665
samples_2 <- 4022
sd_height_2 <- 0.1029

# Active
mean_height_2a <- 1.7032
samples_2a <- 8342
sd_height_2a <- 0.1033

z_2 = 1.96
#Non Active
upper_non_active <- mean_height_2 + z*(sd_height_2/sqrt(samples_2))

lower_non_active <- mean_height_2 - z*(sd_height_2/sqrt(samples_2))

c("Non-active heights:", lower_non_active, upper_non_active)


```

11.  Now, a non-inference task: Determine the number of different options there are in the dataset for the `hours_tv_per_school_day` there are.
```{r}
yrbss %>%group_by(hours_tv_per_school_day)%>% summarise(n())
```

12. Come up with a research question evaluating the relationship between height or weight and sleep. Formulate the question in a way that it can be answered using a hypothesis test and/or a confidence interval. Report the statistical results, and also provide an explanation in plain language. Be sure to check all assumptions, state your $\alpha$ level, and conclude in context.

Question: Is there a relationship with height and sleep?
Hypothesis: The average height of students has no affect on the average number of hours of sleep students receive on school nights.

```{r}
unique(yrbss$school_night_hours_sleep)
length(unique(yrbss$school_night_hours_sleep))



```

```{r}
ggplot(yrbss, aes(x = height, y = school_night_hours_sleep)) + geom_boxplot()

```

```{r}
desc <- describeBy(yrbss$height, yrbss$school_night_hours_sleep, mat=TRUE)
desc$Var <- desc$sd^2
print(desc, row.names=FALSE)



```
The sample sizes of each sleeping group are greater than 30 and based on the boxplots and means calculated for each sleeping group, there are very few particularly extreme values. But we can't make a conclusion based on this. 

