---
title: "Assignment 4 - 607"
author: "Sangeetha Sasikumar"
date: "10/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(ggplot2)
```

####Loading in the data from an excel sheet which I converted as csv. The key part of this assignment was to pass the blanks in as "NA".
```{r}
flights<-read.csv("https://raw.githubusercontent.com/Sangeetha-007/R-Practice/master/607/Assignments/Assignment%204/arrival%20delays%20.csv",na.strings=c("","NA"))
print(flights)
```


####I am listing the column names. 
```{r}
colnames(flights)
la <-pull(flights)  #dyplyr 
print(la) #this line of code is just me playing around with dyplyr, it has nothing to do with the actual task!
```

####This code in this chunk is wrong, and thus commented out, but I kept it for me to sort of look back upon and see the wrong code I tried. 
```{r}
#flights_long <- flights %>% gather(Los.Angeles, Phoenix, San.Diego, San.Francisco, Seattle, -c("X", "X.1"))
#flights_long <- flights %>% gather(Los.Angeles -c(X, X.1))
```

####Here, I am pivoting the wide data into long format. I renamed 2 columns, and filled in the "NA" values with a top-down approach to filling data. 
```{r}
flights_long<-pivot_longer(flights, -c(X, X.1), values_to = "Flight_Count", names_to = "Destination")
head(flights_long)
flights_long2 <- fill(flights_long, X, .direction = c("down"))
head(flights_long2)

```

####I dropped the "NA" values and renamed the first 2 columns. 
```{r}
flights_long2<-flights_long2 %>% drop_na() 
colnames(flights_long2)[1] = "Airline"
colnames(flights_long2)[2] = "Time"
print(n=25,flights_long2)
```

####I took a summary of all the data present. 
```{r}
summary(flights_long2)
```

####I calculated the interquartile range (IQR) of `Flight_Count`. The IQR measures the spread of the middle half of the data.
```{r}
print(IQR(flights_long2$`Flight_Count`))
```
####I went on to calculate IQR and median grouped by airline. 
```{r}
flights_long2 %>%
  group_by(Airline) %>%
  summarise(median_arrivalDelay = median(`Flight_Count`), iqr_arrivalDelay = IQR(`Flight_Count`))
```

####Then for each destination I analyzed it's summary. 
```{r}
angeles<-flights_long2%>%group_by(Destination)%>%filter(Destination=="Los.Angeles")
summary(angeles)
```

```{r}
phoenix<-flights_long2%>%group_by(Destination)%>%filter(Destination=="Phoenix")
summary(phoenix)
```

```{r}
diego<-flights_long2%>%group_by(Destination)%>%filter(Destination=="San.Diego")
summary(diego)
```

```{r}
francisco<-flights_long2%>%group_by(Destination)%>%filter(Destination=="San.Francisco")
summary(francisco)
```

```{r}
seattle <-flights_long2%>%group_by(Destination)%>%filter(Destination=="Seattle")
summary(seattle)
```

####Finally, I made bar graphs filtering the data between "on time" and "delayed". 

```{r}
onTime<-flights_long2%>%filter(Time=="on time")
onTime

ggplot(onTime, aes(fill=`Airline`, y=`Flight_Count`, x=Destination)) + 
    geom_bar(position="dodge", stat="identity")
```

```{r}
delayed<-flights_long2%>%filter(Time=="delayed")
delayed

ggplot(delayed, aes(fill=`Airline`, y=`Flight_Count`, x=Destination)) + 
    geom_bar(position="dodge", stat="identity")
```
