---
title: "Project 2- 607, Part 2"
author: "Sangeetha Sasikumar"
date: "10/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyr)
library(dplyr)
options(dplyr.summarise.inform = FALSE)
```
##Loaded the csv.
```{r}
grades<-read.csv("https://gist.githubusercontent.com/Kimmirikwa/b69d0ea134820ea52f8481991ffae93e/raw/4db7b1698035ee29885d10e1a59bd902716ae168/student_results.csv")
grades
```


##Split the column where the data was appearing with the sex and age concatenated with a "_". I split that data into 2 columns: Gender and Age. 
```{r }
class(grades)

grades[c('Gender', 'Age')] <- str_split_fixed(grades$sex.and.age, '_', 2)

grades 
```
##Then I removed the sex.and.age column. 
```{r}
grades

grades <- subset(grades, select = -c(4) )

grades
```
##Next, I gathered columns 5 to 7, to fill up the columns "Term" and "Grades". 
```{r}
grades<- grades %>%
    gather("Term", "Grades", 5:7)
grades

```
##Next, I wanted to break up the "Term" column, by the dot, so I found out about cbind. It allowed me to create a new dataframe called "new". 
```{r}
new <- cbind(read.table(text = as.character(grades$Term), sep=".", 
         header = FALSE, col.names = c("Col1", "Col2")))
new

class(new)
```
##I used cbind to paste the new dataframe, and the previous dataframe (grades) to paste them together. 
```{r}
grades_updated <- cbind(grades, new)
grades_updated

colnames(grades_updated)
```
##Next, I got rid of the Col1 column and the Term column. 
```{r}
grades_updated <- subset(grades_updated, select = -c(Col1, Term))
grades_updated
```
##Renamed Col2 to Term. 
```{r}
grades_updated <- grades_updated %>%
  rename(Term = Col2)

grades_updated
```
##I needed "test.number"'s values to become variable names. 7 is the column where values will fill in under the new variables created. 
```{r}
grades_updated<-spread(grades_updated, "test.number", 7)
grades_updated
```
##I wanted to get the mean grade, grouped by the phone number (average of each student's test 1).
```{r}
grades_updated %>% 
      group_by(phone) %>% 
      summarise(Mean=mean(`test 1`))

```
##Next, I took the average of each student's Test 1 and Test 2 throughout each term. 
```{r}
averagesOfeachStudent<- grades_updated %>% 
      group_by(name, Term) %>% 
      summarise(Average=(`test 1`+`test 2`)/2)

averagesOfeachStudent
```

##Finally, I took the average of each student based on both tests from all 3 terms. 
```{r}
averagesOfeachStudent %>% group_by(name)%>% 
      summarise(Mean=mean(Average))
```
Sources: 
https://www.listendata.com/2015/06/r-keep-drop-columns-from-data-frame.html,
https://www.geeksforgeeks.org/how-to-split-column-into-multiple-columns-in-r-dataframe/, 
https://stackoverflow.com/questions/66953570/r-split-one-column-into-two-when-the-divider-is-a-dot,
https://www.programmingr.com/examples/r-dataframe/cbind-in-r/,
https://stackoverflow.com/questions/4605206/drop-data-frame-columns-by-name,
http://statseducation.com/Introduction-to-R/modules/tidy%20data/spread/





