---
title: "Assignment3-607"
author: "Sangeetha Sasikumar"
date: "9/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```


1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"
```{r }
majors<-read.csv(paste0("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"))
head(majors)

```

```{r}
#str(majors)
pattern<-'DATA|STATISTICS'
grep(pattern, majors$Major, value=TRUE,ignore.case = TRUE)
#case is ignored because we cant always know if the data is lower case or upper case

```

2. Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
fruitsAndVeggies <- ' "bell pepper" "bilberry"  "blackberry" "blood orange" "blueberry"    "cantaloupe"   "chili pepper" "cloudberry" "elderberry"  "lime" "lychee"  "mulberry"  "olive"  "salal berry"'
fruitsAndVeggies

#new_df<-str_extract_all(fruitsAndVeggies, "\\b[a-z]+\\b")
#new_df
#str_c works, you need to imagine that you are building up a matrix of strings. Each input argument forms a column, and is expanded to the length of the longest argument, using the usual recyling rules.

#new_df[0:1]

extracted <- str_extract_all(fruitsAndVeggies, "\\w[a-z]+\\s?[a-z]+\\w")
print(unlist(extracted))   
class(unlist(extracted))
```

3. Describe, in words, what these expressions will match: \n \n
  * (.)\1\1   : Any character 3 times in a row.
  * "(.)(.)\\2\\1"   : 2 characters that repeat immediately in the reverse order.
  * (..)\1   : 2 characters that repeat immediately in the same order.
  * "(.).\\1.\\1"   : Single character that repeats 2 more times, with each repetition after another single character. 
  * "(.)(.)(.).*\\3\\2\\1"   : Any 3 characters that repeat in the reverse order after any number variable characters.

4. Construct regular expressions to match words that:
* Start and end with the same character.    `^([a-z]).*\1$`
* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)    `([a-zA-Z][a-zA-Z]).*\1`
* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.) - `([a-zA-Z]).*\1.*\1` 

```{r}
#Testing: (I ran out of time to do this better)
mylist<-list("anna", "aann", "church", "eleven")
output_1 <- str_extract_all(mylist, "^([a-z]).*\1$")
print(output_1)
output_2<-str_extract_all(mylist, "([a-zA-Z][a-zA-Z]).*\1")
print(output_2)
output_3<-str_extract_all(mylist, "([a-zA-Z]).*\1.*\1")
print(output_3)

```
